---
title: "Data analysis Casasanto 2009's replication"
author: "Fernando Marmolejo-Ramos"
date: "04/09/2022"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

The description of the rationale behind the following analyses is provided in the manuscript; hence, not much would be elaborated here.

The following model is analysed via GLMM with the Binomial distribution as the dependent variable is dichotomous:

$P \sim a + g + h + v + s + (1|c/p)$

The six independent variables were age (a), gender (g; 2 levels; male and female), FLANDERS handedness score (h), story version (v; 4 levels; 4 versions), and survey modality (s; 2 levels; online and paper-and-pencil). Participants (p) and country (c; nine levels; the sites where data were collected) were modelled as random effects. `(1|c/p)` indicates that participants are nested within countries (numeric variables are italicized and no interactions were considered). 


First off, let's read the data:

```{r, warning=FALSE, message=FALSE}

library(readxl)
data <- read_excel(file.choose())
dim(data)

# retaining complete cases only
data.complete <- data[complete.cases(data), ]
dim(data.complete)

# checking the data
library(skimr)
skim(data.complete)

# second check
apply(is.na(data.complete), 2, which)


# checking the data's structure
str(data.complete)

# transforming variables:
data.complete <- transform(data.complete,
                 participant_ID = as.factor(participant_ID),
                 age = as.numeric(age),
                 gender = as.factor(gender),
                 quest_ver = as.factor(quest_ver),
                 lab = as.factor(lab),
                 good_right = as.factor(good_right),
                 modality = as.factor(modality)
                           )

str(data.complete)
apply(is.na(data.complete), 2, which)


# there were 8 versions in the japanese sample so such data will be excluded

# also, retain males and females only

data.complete2 <- subset(data.complete, gender=='m' | gender =='f')

data.complete2 <- subset(data.complete2, quest_ver=='1'|quest_ver=='2'|quest_ver=='3'| quest_ver=='4')

str(data.complete2)

# note unused levels remain; so let's drop them...

data.complete2$quest_ver <- droplevels(data.complete2$quest_ver)

data.complete2$gender <- droplevels(data.complete2$gender)

str(data.complete2)
dim(data.complete2)

skim(data.complete2)

# second check
apply(is.na(data.complete2), 2, which)
```


# Statistical modelling

```{r, warning=FALSE, message=FALSE}

library(lme4)
model <- glmer(good_right ~ age + gender + hand_score +
                 quest_ver + modality + 
                 (1|lab/participant_ID), 
               data=data.complete2,
              family=binomial)

summary(model, corr=F)

# odds ratios for 'hand_score'
exp(summary(model)$coefficients[4,1])
exp(confint(model, parm='hand_score'))

# ANOVA table for the model
library(car)
Anova(model)

# model's performance
library(performance)

model_performance(model)

```

The model suggested main effects of `handedness_score` and `question_version` only.

_FLANDERS handedness score_: Each one-unit change in `handedness_score` will increase the log odds of choosing 'right-is-good' by 0.043, and its p-value indicates that it is significant in choosing 'right-is-good'. In an odds-ratio jargon, it can be said that the odds of choosing 'right-is-good' increase by a factor of 1.04.


The variance explained by the full model was about 10% and the variance and the variance explained by the fixed effects only was about 7%. AIC=2647 and BIC=2647

## Simplified model

A second (reduced) model could be considered such that only significant variables are retained. Note too an interaction is added to investigate if there's any:

```{r, warning=FALSE, message=FALSE}

model.simple <- glmer(good_right ~ hand_score *
                 quest_ver +  
                 (1|lab/participant_ID), 
               data=data.complete2,
              family=binomial)

summary(model.simple, corr=F)

# odds ratios for 'hand_score'
exp(summary(model.simple)$coefficients[2,1])
exp(confint(model.simple, parm='hand_score'))

# ANOVA table for the model
Anova(model.simple)


# model's performance
model_performance(model.simple)
```

The model suggested no interaction and the main effects of `handedness_score` and `question_version` remained.

_FLANDERS handedness score_: Each one-unit change in `handedness_score` will increase the log odds of choosing 'right-is-good' by 0.051, and its p-value indicates that it is significant in choosing 'right-is-good'. In an odds-ratio jargon, it can be said that the odds of choosing 'right-is-good' increase by a factor of 1.05.

The variance explained by the full model was about 9.8% and the variance and the variance explained by the fixed effects only was about 6.9%. AIC=2654 and BIC=2655.



## On optimisers

It's known that in some cases there are convergence issues. In case this occurs, here's is a work around (taken from \url{https://joshua-nugent.github.io/allFit/}):

```{r, warning=FALSE, message=FALSE}
library(optimx)
library(dfoptim)
library(parallel)

ncores <- detectCores()
diff_optims <- allFit(model, maxfun = 1e5, parallel = 'multicore', ncpus = ncores)
```


# Spineplot

While the effect of `question_version` suggests some versions of the task led to effects on the classification of 'right-is-good', the true interest is in the handedness score and its effect on the potential association 'good-right' and 'good-left'.

This association is visualized via a conditional density plot:

```{r, warning=FALSE, message=FALSE}

spine.plot <- with(data.complete2,
spineplot(good_right ~ hand_score,
          xlab = 'FLANDERS handedness score',
          ylab='Horizontal allocation (1=right is good, 
          \n0=left is good)',
          col=c('grey90','grey35')))


# a Conditional Density Plots can be used but its result isn't too clear to understand

#with(data.complete2,
#     cdplot(good_right ~ hand_score)
#     )

# check also the 'vcd' package for the function 'spine'

# to obtain actual counts and proportions try...
prop.table(spine.plot)*100
```


# Supplementary analysis

See rationale in the manuscript.

## Classification algorithms

### One rule

```{r, warning=FALSE, message=FALSE}
library(OneR)

model1 <- 
  with(data.complete2, 
  as.formula(good_right ~ age + gender + hand_score +
                 quest_ver + modality + 
                 lab))


# 1R approach
cat.rank <- OneR(model1, 
                 verbose=T, data = data.complete2)
cat.rank
summary(cat.rank)
plot(cat.rank)
```

The 1R algorithm lists both `hand_score` and `quest_ver` in the first place. Both with a classification accuracy of 60.84%. `lab` occupied the 3rd place with a classification accuracy of 59.97%.

### Boruta

```{r, warning=FALSE, message=FALSE}
library(Boruta)
set.seed(2022)

boruta.selection <- Boruta(model1, 
                 data = data.complete2)

print(boruta.selection)

plot(boruta.selection)
# Recall: green are confirmed variables 
#         red are rejected variables
#         blue are shadow variables
# The Y axis represents level of importance

getConfirmedFormula(boruta.selection)

ranking.variables <- attStats(boruta.selection)

ranking.variables[order(ranking.variables$medianImp, 
                        decreasing = T),]
```

The Boruta algorithm ranks the importance of variables, in this order, as `quest_ver`, `lab` and `hand_score`.

The two common variables were `quest_ver` and `hand_score`. A logistic model considering these two variables was already examined in the section 'simplified model' and therefore will not be pursued further.

# Overall message:

* The likelihood of choosing 'right-is-good' increases as the score in the FLANDERS handedness test increases.

* However, that likelihood is not strikingly large (nearing 1).




